Mission-time LTL (MLTL) Formula Validation Via Regular Expressions\thanks{Work supported in part by [omitted for blind review] {\bf Note that all references to appendices will be replaced with URLs in the final paper; they are appendices only for blind review.}}

### 2022-10-13 paper/Paper.tex
vector->set + other wording changes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{-0.1in}
\section{MLTL Regular Expressions} \label{regex}
\vspace{-0.1in}
We extend the standard definition of Regular Expressions (regex) \cite{sipser} to introduce notation that describes the satisfying computations of an MLTL formula.
\begin{definition}[\emph{Temporal Regular Expression}] \label{TRE} 
Let $R$ and $T$ denote regular expressions, and let $S$ be an abbreviation for $(0 \ | \ 1)$. Let fixed $n \in \mathbb{N}$ denote the number of propositional variables in a MLTL formula. We use the following operations to describe the form of satisfying computations of the formula in the bit string representation:
\begin{itemize}
  \item \emph{(Concatenation)} $RT$ describes the set of strings obtained by concatenating any string generated by $R$ and any string generated by $T$ in that order.
 \item \emph{(Alternation)} $R \lor T$ describes the set union of the strings generated by $R$ and $T$.
    \item Pad($R$, $T$) determines which regular expression is longer and concatenate $(,S^n)$ repeatedly to the end of the shorter regular expression until the two regular expressions are the same length. Note that in the bit string representation, $(,S^n)$, denotes a time step in which the truth value of all $n$ propositional variables does not matter.
    \item For $R \land T$, we first use Pad($R$, $T$), and then take the set intersection of the sets of strings described by the two regular expressions.
      \item $R^i$ denotes regular expression consisting of $R$ repeated $i$ times for $i \geq 0$. $R^0 = \epsilon$. %KYR: is this one relevant or should we take it out?
    \item We do not use the Kleene star since our computations are a fixed finite length.
\end{itemize}
\end{definition}
%An example of these operations follows.
\begin{example} %KYR: NICE example!
Let $n = 2$, and let $R = S1$ and $T = 1S,1S | S1,S1$.\\
To compute $R \land T$ and $R \lor T$, we use Pad($R$,$T$). Since $T$ is the longer regex by one time step, we concatenate ``$,S^2$'' to the right of $R$. Thus $T =1S,1S | S1,S1$ and $R = S1,SS$. Now the regexes are the same length, so we can perform set intersection and alternation:
\begin{itemize}
    \item $R \land T = 11,1S \lor S1, S1$.
    \item $R \lor T = S1,SS \lor 1S,1S \lor S1,S1 = S1,SS \lor 1S,1S$.
\end{itemize}
\end{example}

@@

\section{WEST Algorithm and Analysis} \label{WESTalg}
We now present the  full WEST algorithm, which recursively computes all satisfying temporal regular expressions to a given MLTL formula (in NNF). 
We use C++ vectors of strings to represent alternation of regular expressions; for instance $\{t_1, t_2, ..., t_n\} = \bigcup_{i=1}^n t_i$. 
Additionally, we provide details for performing set intersection of temporal regular expressions, and the algorithm for the temporal operator $\mathcal{F}$. 
Note how the algorithm parallels the regular expression defined for $\mathcal{F}$, and the algorithms for the other three temporal operators follow an identical structure.

One major detail that drastically improves running time is applying a simplifying function to the output of set intersection. 
This is because the set intersection operation is repeatedly used to compute the temporal operators $\mathcal{G}, \mathcal{U}, \mathcal{R}$. 
Although repeated set intersection blows up exponentially in space in the worse case, it rarely does in practice; thus why we frequently simplify outputs. 
We simplify alternation by using the fact that for regular expressions $w_1$ and $w_2$, $\mathscr{L}(w_1 1 w_2 \lor w_1 0 w_2) = \mathscr{L}(w_1 s w_2)$, and check if the reduction applies at every index between every pair of strings in a given vector.

@@

% ZILI: Presenting algorithm vertically instead
\hspace{-5mm}
\begin{minipage}{0.52\textwidth}
\begin{algorithm}[H]
\caption{WEST Algorithm \\
Inputs: $\phi$ - MLTL formula in NNF \\
$\phi_1$ and $\phi_2$ below are subformulas of $\phi$\\
$n$ - number of propositional variables \\
Output: Vector of REGEX satsfying $\phi$}
\begin{algorithmic}[1]
    \Procedure{reg}{string $\phi$, int $n$}
        \If{$\phi$ is $\top$ or $\bot$}
            \State return reg\_prop\_const($\phi$, n)
        \EndIf

@@

        \If{$\phi = \phi_1 \mathcal{R}_{[a, b]} \phi_2$}
            \State return reg\_R(reg($\phi_1$), reg($\phi_2$), a, b, n)
        \EndIf
    \EndProcedure
\end{algorithmic}
\end{algorithm}
\vspace{5.9mm}
\end{minipage}
\hfill
\begin{minipage}{0.48\textwidth}
\begin{algorithm}[H]
\caption{set\_intersect}
Inputs: $R, T$ - Two Vectors of REGEX\\
$n$ - number of propositional variables\\
Output: Vector equal to $R \land T$
\begin{algorithmic}[1]
\Procedure{set\_intersect}{$R$, $T$, n}
    \State Pad($R$, $T$, n), return $\leftarrow \{\}$
    \For{$(r, t) \in R\times T$}
            \State add bit\_wise\_and($r, t$) to ret 
    \EndFor
    \State return simplify(ret)
\EndProcedure
\end{algorithmic}
\end{algorithm}
\vspace{-6.5mm}
\begin{algorithm}[H]
\caption{reg\_U} 
Inputs: r\_phi, r\_psi - vectors of REGEX for MLTL formulas $\phi$ and $\psi$ \\
$a, b$ - integers representing interval bound \\
$n$ - number of propositional variables\\
Output: Vector of REGEX for $\phi \mathcal{U}_{[a, b]} \psi$
\begin{algorithmic}[1]
\Procedure{reg\_U}{r\_phi, r\_psi, $a$, $b$, $n$} 
    \State comp $\leftarrow$ ((`$S$')$^{n}$ + `,')$^{a} + $ r\_psi
    \If{$a > b$} return $\{\}$
    \EndIf
    \For{($a \leq i \leq b-1$) }
        \State temp\_G1 $\leftarrow$ reg\_G(r\_phi, $a$, $i$, $n$)
        \State \parbox[t]{.8\linewidth}{temp\_G2 $\leftarrow$ reg\_G(r\_psi, $i+1$, \hspace*{5mm}$i+1$, $n$)}
        \State \parbox[t]{.8\linewidth}{temp\_comp $\leftarrow$ set\_intersect( \hspace*{5mm}temp\_G1, temp\_G2, n)}
        \State \parbox[t]{.8\linewidth}{comp $\leftarrow$ join( comp, temp\_comp)}
        % , \hspace*{5mm}simp)}
    \EndFor
    \State return comp
\EndProcedure
\end{algorithmic}
\end{algorithm}
\end{minipage}

\subsection{Theoretical Correctness of the WEST Algorithm}
Correctness of the WEST algorithm is dependent on the correctness of subroutines \textit{reg\_prop\_cons}, \textit{reg\_prop\_var}, \textit{join}, \textit{set\_intersect}, \textit{reg\_F}, \textit{reg\_G}, \textit{reg\_U}, and \textit{reg\_R}. 
The routines \textit{reg\_prop\_cons} and \textit{reg\_prop\_var} take as input a MLTL formula of the appropriate form, and return the regular expression as %previously
defined in section \ref{regex}.
\textit{Join} concatenates two input vectors of computation strings $R$ and $T$, which is equivalent to $\mathscr{L}(R) \cup \mathscr{L}(T)$. 
\textit{Set\_intersect} takes as input two vectors of computation strings $R$ and $T$, where $R$ represents $r_1 | ... | r_a$ and $T$ represents $t_1 | ... | t_b$ such that each $r_i$ and $t_j$ are regular expressions over $\Sigma = \{`0', `1', `S', `,'\}$.
Assume that all strings of regular expressions are right padded to equal length. 
We show that $\mathscr{L}(\textit{set\_intersect}(R, T)) = \mathscr{L}(R) \cap \mathscr{L}(T)$. Since a vector of regular expressions is synonymous to the collective \texttt{OR} of each of them, we have that 
\begin{align*}
    \mathscr{L}(R) \cap \mathscr{L}(T) &= \left(\bigcup_{i = 0}^a \mathscr{L}(r_i) \right)\cap \left( \bigcup_{j = 0}^b \mathscr{L}(t_j)\right) = \bigcup_{i = 0}^a \bigcup_{j = 0}^b \left(\mathscr{L}(r_i) \cap \mathscr{L}(t_j)\right).
\end{align*}
The for loop in line 3 of \textit{set\_intersect} computes the union of \textit{bit\_wise\_and}$(r_i, t_j)$ over all such pairs, and so it suffices to show $\mathscr{L}(\textit{bit\_wise\_and}(r_i, t_j)) = \mathscr{L}(r_i) \cap \mathscr{L}(t_j)$.  
Given a computation $\pi$, $\pi \in \mathscr{L}(r_i) \cap \mathscr{L}(t_j)$ if and only if $\pi$ matches every character of both $r_i$ and $t_j$. 
$\textit{Bit\_wise\_and}(r_i, t_j)$ compares $r_i$ and $t_j$ character by character and computes their intersection, which is defined naturally: $0 \cap 1 = \emptyset,\ 0 \cap S = 0,\ 1 \cap S = 1, \ 0 \cap 0 = 0, \ 1 \cap 1= 1,\text{ and } S \cap S = S$. 
Note that this operation is commutative. This exhaustively captures all the cases for which $\pi$ must match corresponding characters from $r_i$ and $t_j$. 
Thus $\mathscr{L}(\textit{bit\_wise\_and}(r_i, t_j)) = \mathscr{L}(r_i) \cap \mathscr{L}(t_j)$ and the claim holds.

@@

\hspace{3mm} If $\phi$ is a propositional constant or variable, it's easy to see that $S(\text{\emph{reg}}(\bot)) = 0$ since only the empty vector is returned. By definition $\text{\emph{reg}}(\top) = S^n$, so we have that $S(\text{\emph{reg}}(\top)) = O(n)$. Similarly, $\text{\emph{reg}}(p_k)$ and $\text{\emph{reg}}(\neg p_k)$ both return strings of the same length and thus $S(\text{\emph{reg}}(p_k)) = S(\text{\emph{reg}}(\neg p_k)) = S(\text{\emph{reg}}(\top)) = n$.\\
\hspace{3mm} If $\phi$ is ``$\phi_{1} \lor \phi_{2}$'' we return $\textit{join}(\text{\emph{reg}}(\phi_{1}), \text{\emph{reg}}(\phi_{2}))$, which simply concatenates the contents of the two vectors. 

@@

$ \textit{set\_intersect}(\text{\emph{reg}}(\phi_{1}), \text{\emph{reg}}(\phi_{2}), n)$ returns a vector of size $S(A) \cdot S(B)$ in the worst case when no simplification can be made. 

@@

In this calculation, $(n+1)i$ counts the concatenation of the padded component of the computation in the worse case when we have a vector of only length 1 strings. \\

@@

If in the worst case no simplification occurs in any call of \textit{set\_intersect}, space complexity remains unchanged, but simplifying a vector of regular expressions is cubic in input size. In practice however, both time and space complexities are much more optimistic than worst-case estimates.

@@

We accompany our proof of algorithmic correctness with a rigourous evaluation of implementation correctness: that our WEST tool correctly implements our WEST algorithm. % The correctness of an abstract algorithm one can describe on pencil and paper differs from the correctness of an actual implementation of the algorithm (the program). 

@@

Since the primary purpose of the WEST tool is valiation of formulas for humans, we focused on writing a tool with an intutiive human-facing interface then verifying that implementation rather than wrestling with verified code generated by a tool made to optimize for different objectives.

One naive approach involves testing every possible input up to a certain size to verify the output or behaviors of the program.
For WEST, this strategy would generate a very large test suite with signficant redundancy. 
For instance, there is little sense in testing all MLTL formulas of the form $p_0 \mathcal{U}_{[0,t]} p_1$ for all $t$ such that $0 \leq t \leq 99$; verification of a few should give sufficient confidence of correctness of the program. 
We approach this problem more intelligently: we test our implementation with a test suite that explores all control paths (up to a certain depth) of the algorithm. A control path is a sequence of lines of code that are executed in one calculation of an input formula. Differing structures of formulas will lead to different sequences of lines of code to be executed, and a test suite to explore all such paths provides high confidence in the correctness of the WEST tool.

@@

The substrings ``$0$", ``$1S0$", and ``$1$" are inserted at the beginning, middle, and end, respectively. In general, a vector of length $m$ strings seeking to apply the theorem to $n$ of its columns has $m \choose n$ combinations of columns to check for. Implementation of this algorithm can drastically increase run-time while proving its utility in only a small number of circumstances.

@@

Theorem \ref{simp} addresses a non-trivial situation in which a vector of regular expressions may be simplified. A natural question to ask is what is the minimum number of strings of regular expressions needed to represent a language of computations. However, such a minimal representation is not unique. For instance, $\mathscr{L}(S1|1S) = \mathscr{L}(S1|10) = \{01, 10, 11\}$. There may be other schemes needed to simplify any arbitrary union of regular expressions to their minimal representation.\\

@@

To compute the satisfying computations of an MLTL formula, many of the functions in the WEST program require the regular expressions of the satisfying computations of the subformulas as inputs. We denote these regexes by $R$ and $T$, where $R = R_1 | R_2 | ... | R_k$ is represented as a vector of strings $R = {R_1, R_2, ..., R_k}$, and $T = T_1 | T_2 | ... | T_\ell$ is represented as a vector of strings $T = {T_1, T_2, ..., T_\ell}$. Additionally, $n$ will always refer to the number of propositional variables, and nnf refers to an input formula in negation normal form.
\begin{algorithm}[H]
\caption{Pad a vector to elements of all equal length}
%Input: Vector of strings that represents a regex, number of propositional variables \\
%Output: Vector of strings padded to equal length
\begin{algorithmic}
\Procedure{Pad}{vector $R$, int $n$} 
    \State max\_Length $\leftarrow$ $\max_{\{r \in R\}}$($r$.length())
    \For{($r \in R$)}
        \State diff $\leftarrow$ (max\_length$- r$.length()) / $(n+1)$ 
        \State  $r \leftarrow r + (,S^n)^{\text{diff}}$
    \EndFor
    \State return $R$

\EndProcedure
\end{algorithmic}
\end{algorithm}
\begin{algorithm}[H]
\caption{Computes regex for propositional constant}
%Input: String that is either ``T" or ``!", number of propositional variables\\
%Output: Vector of strings that represents the appropriate satisfying computations
\begin{algorithmic}
\Procedure{reg\_prop\_cons}{string nnf, int $n$}
    \If {(nnf $=$ ``T" and $n \neq 0$)} return $\{S^n\}$
    \Else \hspace{1mm} return $\{\}$ 
    \EndIf
\EndProcedure
\end{algorithmic}
\end{algorithm}
\begin{algorithm}[H]
\caption{Output the vector of computation satisfying a propositional variable.}
%Input: String that represents a propositional variable or the negation of one, number of propositional variables \\
%Output: Vector of strings that represents the appropriate satisfying computations
\begin{algorithmic}
\Procedure{reg\_prop\_var}{string nnf, int $n$}
    \If {(nnf $=$ ``p$k$", where $k$ is a nonnegative integer)} return $\{S^k1S^{n-k-1}\}$
    \EndIf 
    \If {(nnf $=$ ``$\scriptstyle{\sim}$p$k$", where $k$ is a nonnegative integer)} return $\{S^k0S^{n-k-1}\}$
    \EndIf
\EndProcedure
\end{algorithmic}
\end{algorithm}
\begin{algorithm}[H]
\caption{Combines two strings that differ only by one character into one}
%Input: Two strings that represent regexes\\
%Output: Single string that represents computations represented by both input strings or FAIL
\begin{algorithmic}
\Procedure{simplify\_string}{string $r$, string $t$}   
    \If{($r$.length() $\neq$ $t$.length())} exit
        \EndIf
    \For{($0 \leq i < r$.length())}
        \State pre\_r $\leftarrow r[0, i-1]$ 
        \State char\_r $\leftarrow r[i]$ 
        \State post\_r $\leftarrow r[i+1, r\text{.length()}-1]$ 
        \State pre\_t $\leftarrow t[0, i-1]$ 
        \State char\_t $\leftarrow t[i]$ 
        \State post\_t $\leftarrow t[i+1, t\text{.length()}-1]$
        \If{(pre\_r $=$ pre\_t and post\_r $=$ post\_t)}
            \If{(char\_r $\neq$ char\_t)}
                \State return pre\_r +``$S$" + post\_r 
            \Else
                \State return $r$
            \EndIf
        \EndIf
    \EndFor
    \State return FAIL 
\EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
\caption{Simplifies a vector of strings using simplify\_string}
%Input: Vector of strings representing a regex\\
%Output: Vector of strings representing a regex simplified using simplify\_string
\begin{algorithmic}
\Procedure{simplify}{vector $R$}    
    \State Pad all strings in $R$ to the same length
    \If{($R$.length() $\leq 1$)} return $R$
        \EndIf
    \State $i = R\text{.length} - 1$
    \State $j = i - 1$

@@

\caption{Takes union of two regexes (combines two vectors into one)}
%Input: Vectors of strings that represent regexes, simplify boolean\\
%Output: Vector of strings that represents union of inputted regexes

@@

\Procedure{join}{vector $R$, vector $T$, bool simp}

@@

% Inputs: Vector of strings representing the regex for $\phi$, interval bounds, number of propositional variables, simplify boolean\\
% Output: Vector of strings that represents the appropriate satisfying computations 

@@

\Procedure{reg\_F}{vector reg\_phi, int $a$, int $b$, int $n$, bool simp} 

@@

%Inputs: Vector of strings representing the regex for $\phi$, interval bounds, number of propositional variables, simplify boolean\\
%Output: Vector of strings that represents the appropriate satisfying computations

@@

\Procedure{reg\_G}{vector reg\_phi, int $a$, int $b$, int $n$, bool simp} 

@@

%Inputs: Vectors of strings representing the regexes for $\phi$ and $\psi$, interval bounds, number of propositional variables, simplify boolean\\

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{-0.1in}
\section{MLTL Regular Expressions} \label{regex}
\vspace{-0.1in}
We extend the standard definition of Regular Expressions (regex) \cite{sipser} to introduce notation that describes the satisfying computations of an MLTL formula.
\begin{definition}[\emph{Temporal Regular Expression}] \label{TRE} 
Let $R$ and $T$ denote regular expressions, and let $S$ be an abbreviation for $(0 \ | \ 1)$.\\
Let fixed $n \in \mathbb{N}$ denote the number of propositional variables in a MLTL formula. We use the following operations to describe the form of satisfying computations of the formula in the bit string representation:
\begin{itemize}
  \item \emph{(Concatenation)} $RT$ describes the set of strings obtained by concatenating any string generated by $R$ and any string generated by $T$ in that order.
 \item \emph{(Alternation)} $R \lor T$ describes the set union of the strings generated by $R$ and $T$.
\item Pad($R$, $T$) determines which regular expression is longer and concatenate $(,S^n)$ repeatedly to the end of the shorter regular expression until the two regular expressions are the same length. Note that in the bit string representation, $(,S^n)$, denotes a time step in which the truth value of all $n$ propositional variables does not matter.
\item For $R \land T$, we first use Pad($R$, $T$), and then take the set intersection of the sets of strings described by the two regular expressions.
\item $R^i$ denotes regular expression consisting of $R$ repeated $i$ times for $i \geq 0$. $R^0 = \epsilon$. %KYR: is this one relevant or should we take it out?
\item We do not use the Kleene star since our computations are a fixed finite length.
\end{itemize}
\end{definition}
%An example of these operations follows.
\begin{example} %KYR: NICE example!
Let $n = 2$, and let $R = S1$ and $T = 1S,1S | S1,S1$.
To compute $R \land T$ and $R \lor T$, we use Pad($R$,$T$). Since $T$ is the longer regex by one time step, we concatenate ``$,S^2$'' to the right of $R$. Thus $T =1S,1S | S1,S1$ and $R = S1,SS$. Now the regexes are the same length, so we can perform set intersection and alternation:
\begin{itemize}
    \item $R \land T = 11,1S \lor S1, S1$.
    \item $R \lor T = S1,SS \lor 1S,1S \lor S1,S1 = S1,SS \lor 1S,1S$.
\end{itemize}
\end{example}

@@

\section{WEST Algorithm and Analysis} \label{WESTalg}
We now present the  full WEST algorithm, which recursively computes all satisfying temporal regular expressions to a given MLTL formula (in NNF). 
For convenience, we use sets to represent alternation of regular expressions; for $n$ regular expressions $t_0, ..., t_{n-1}$, we'll write $\{t_1, t_2, ..., t_n\} = \bigcup_{i=1}^n t_i$. 
Additionally, we provide details for performing set intersection of temporal regular expressions, and the algorithm for the temporal operator $\mathcal{F}$. 
Note how reg\_U parallels the regular expression defined for $\mathcal{F}$, and the algorithms for the other three temporal operators follow an identical structure.

One major detail that drastically improves running time is applying a simplifying function to the output of set intersection. 
This is because the set intersection operation is repeatedly used to compute the temporal operators $\mathcal{G}, \mathcal{U}, \mathcal{R}$. 
Although repeated set intersection blows up exponentially in space in the worse case, it rarely does in practice; thus why we frequently simplify outputs. 
We simplify alternation by using the fact that for regular expressions $w_0$ and $w_1$, $\mathscr{L}(w_0 1 w_1 \lor w_0 0 w_1) = \mathscr{L}(w_0 s w_1)$, and check if the reduction applies at every index between every pair of strings in a given set.

@@

% ZILI: Presenting algorithm vertically instead
\hspace{-5mm}
\begin{minipage}{0.52\textwidth} 
\begin{algorithm}[H] 
\caption{WEST Algorithm \\
Inputs: $\phi$ - MLTL formula in NNF \\
$\phi_1$ and $\phi_2$ below are subformulas of $\phi$\\
$n$ - number of propositional variables \\
Output: set of REGEX satisfying $\phi$}
\begin{algorithmic}[1]
    \Procedure{reg}{string $\phi$, int $n$}
        \If{$\phi$ is $\top$ or $\bot$}
            \State return reg\_prop\_const($\phi$, n)
        \EndIf

@@

        \If{$\phi = \phi_1 \mathcal{R}_{[a, b]} \phi_2$}
            \State return reg\_R(reg($\phi_1$), reg($\phi_2$), a, b, n)
        \EndIf
    \EndProcedure
\end{algorithmic}
\end{algorithm}
\vspace{5.9mm}
\end{minipage}
\hfill
\begin{minipage}{0.48\textwidth}
\begin{algorithm}[H]
\caption{set\_intersect}
Inputs: $R, T$ - two sets of REGEX\\
$n$ - number of propositional variables\\
Output: set of REGEX equal to $R \land T$
\begin{algorithmic}[1]
\Procedure{set\_intersect}{$R$, $T$, n}
    \State Pad($R$, $T$, n), return $\leftarrow \{\}$
    \For{$(r, t) \in R\times T$}
            \State add bit\_wise\_and($r, t$) to ret 
    \EndFor
    \State return simplify(ret)
\EndProcedure
\end{algorithmic}
\end{algorithm}
\vspace{-6.5mm}
\begin{algorithm}[H]
\caption{reg\_U} 
Inputs: r\_$\phi$, r\_$\psi$ - sets of REGEX for MLTL formulas $\phi$ and $\psi$ (after calling reg)\\
$a, b$ - integers representing interval bound \\
$n$ - number of propositional variables\\
Output: set of REGEX for $\phi \mathcal{U}_{[a, b]} \psi$
\begin{algorithmic}[1]
\Procedure{reg\_U}{r\_$\phi$, r\_$\psi$, $a$, $b$, $n$} 
    \State comp $\leftarrow$ ((`$S$')$^{n}$ + `,')$^{a} + $ r\_$\psi$
    \If{$a > b$} return $\{\}$
    \EndIf
    \For{($a \leq i \leq b-1$) }
        \State temp\_G1 $\leftarrow$ reg\_G(r\_$\phi$, $a$, $i$, $n$)
        \State \parbox[t]{.8\linewidth}{temp\_G2 $\leftarrow$ reg\_G(r\_$\psi$, $i+1$, \hspace*{5mm}$i+1$, $n$)}
        \State \parbox[t]{.8\linewidth}{temp\_comp $\leftarrow$ set\_intersect( \hspace*{5mm}temp\_G1, temp\_G2, n)}
        \State \parbox[t]{.8\linewidth}{comp $\leftarrow$ join( comp, temp\_comp)}
        % , \hspace*{5mm}simp)}
    \EndFor
    \State return comp
\EndProcedure
\end{algorithmic}
\end{algorithm}
\end{minipage}

\subsection{Theoretical Correctness of the WEST Algorithm}
Correctness of the WEST algorithm is dependent on the correctness of subroutines \textit{reg\_prop\_cons}, \textit{reg\_prop\_var}, \textit{join}, \textit{set\_intersect}, \textit{reg\_F}, \textit{reg\_G}, \textit{reg\_U}, and \textit{reg\_R}. 
The routines \textit{reg\_prop\_cons} and \textit{reg\_prop\_var} take as input a MLTL formula of the appropriate form, and return the regular expression as %previously
defined in section \ref{regex}.
\textit{Join} concatenates two sets of regular expressions $R$ and $T$, which is equivalent to $\mathscr{L}(R) \cup \mathscr{L}(T)$. 
\textit{Set\_intersect} takes as input two sets of regular expressions $R$ and $T$, where $R$ represents $r_1 | ... | r_a$ and $T$ represents $t_1 | ... | t_b$ such that each $r_i$ and $t_j$ are regular expressions over $\Sigma = \{`0', `1', `S', `,'\}$.
Assume that all strings of regular expressions are right padded to equal length. 
We show that $\mathscr{L}(\textit{set\_intersect}(R, T)) = \mathscr{L}(R) \cap \mathscr{L}(T)$:
\begin{align*}
    \mathscr{L}(R) \cap \mathscr{L}(T) &= \left(\bigcup_{i = 0}^a \mathscr{L}(r_i) \right)\cap \left( \bigcup_{j = 0}^b \mathscr{L}(t_j)\right) = \bigcup_{i = 0}^a \bigcup_{j = 0}^b \left(\mathscr{L}(r_i) \cap \mathscr{L}(t_j)\right).
\end{align*}
The for loop in line 3 of \textit{set\_intersect} computes the union of \textit{bit\_wise\_and}$(r_i, t_j)$ over all such pairs, and so it suffices to show $\mathscr{L}(\textit{bit\_wise\_and}(r_i, t_j)) = \mathscr{L}(r_i) \cap \mathscr{L}(t_j)$.  
Given a computation $\pi$, $\pi \in \mathscr{L}(r_i) \cap \mathscr{L}(t_j)$ if and only if $\pi$ matches every character of both $r_i$ and $t_j$. 
$\textit{Bit\_wise\_and}(r_i, t_j)$ compares $r_i$ and $t_j$ character by character and computes their intersection, which is defined naturally: $0 \cap 1 = \emptyset,\ 0 \cap S = 0,\ 1 \cap S = 1, \ 0 \cap 0 = 0, \ 1 \cap 1= 1,\text{ and } S \cap S = S$. 
Note that this operation is commutative. This exhaustively captures all the cases for which $\pi$ must match corresponding characters from $r_i$ and $t_j$. 
Thus $\mathscr{L}(\textit{bit\_wise\_and}(r_i, t_j)) = \mathscr{L}(r_i) \cap \mathscr{L}(t_j)$ and the claim holds.

@@

\hspace{3mm} If $\phi$ is a propositional constant or variable, it's easy to see that $S(\text{\emph{reg}}(\bot)) = 0$ since only the empty set is returned. By definition $\text{\emph{reg}}(\top) = S^n$, so we have that $S(\text{\emph{reg}}(\top)) = O(n)$. Similarly, $\text{\emph{reg}}(p_k)$ and $\text{\emph{reg}}(\neg p_k)$ both return strings of the same length and thus $S(\text{\emph{reg}}(p_k)) = S(\text{\emph{reg}}(\neg p_k)) = S(\text{\emph{reg}}(\top)) = n$.\\
\hspace{3mm} If $\phi$ is ``$\phi_{1} \lor \phi_{2}$'' we return $\textit{join}(\text{\emph{reg}}(\phi_{1}), \text{\emph{reg}}(\phi_{2}))$, which simply concatenates the contents of the two sets. 

@@

$ \textit{set\_intersect}(\text{\emph{reg}}(\phi_{1}), \text{\emph{reg}}(\phi_{2}), n)$ returns a set of size $S(A) \cdot S(B)$ in the worst case when no simplification can be made. 

@@

In this calculation, $(n+1)i$ counts the concatenation of the padded component of the computation in the worse case when we have a set of only length 1 strings. \\

@@

If in the worst case no simplification occurs in any call of \textit{set\_intersect}, space complexity remains unchanged, but simplifying a set of regular expressions is cubic in input size. In practice however, both time and space complexities are much more optimistic than worst-case estimates.

@@

We accompany our proof of algorithmic correctness with a rigorous evaluation of implementation correctness: that our WEST tool correctly implements our WEST algorithm. % The correctness of an abstract algorithm one can describe on pencil and paper differs from the correctness of an actual implementation of the algorithm (the program). 

@@

Since the primary purpose of the WEST tool is validation of formulas for humans, we focused on writing a tool with an intuitive human-facing interface then verifying that implementation rather than wrestling with verified code generated by a tool made to optimize for different objectives.

One naive approach involves testing every possible input up to a certain size to verify the output or behaviors of the program.
For WEST, this strategy would generate a very large test suite with significant redundancy. 
For instance, there is little sense in testing all MLTL formulas of the form $p_0 \mathcal{U}_{[0,t]} p_1$ for all $t$ such that $0 \leq t \leq 99$; verification of a few should give sufficient confidence of correctness of the program. 
We approach this problem more intelligently: we test our implementation with a test suite that explores all control paths (up to a certain depth) of the algorithm. A control path is a sequence of lines of code that are executed in one calculation of an input formula. Differing structures of formulas will lead to different sequences of lines of code to be executed, and a test suite to explore all such paths provides high confidence in the correctness of the WEST tool.

@@

The substrings ``$0$", ``$1S0$", and ``$1$" are inserted at the beginning, middle, and end, respectively. In general, a set of length $m$ strings seeking to apply the theorem to $n$ of its columns has $m \choose n$ combinations of columns to check for. Implementation of this algorithm can drastically increase run-time while proving its utility in only a small number of circumstances.

@@

Theorem \ref{simp} addresses a non-trivial situation in which a set of regular expressions may be simplified. A natural question to ask is what is the minimum number of strings of regular expressions needed to represent a language of computations. However, such a minimal representation is not unique. For instance, $\mathscr{L}(S1|1S) = \mathscr{L}(S1|10) = \{01, 10, 11\}$. There may be other schemes needed to simplify any arbitrary union of regular expressions to their minimal representation.\\

@@

To compute the satisfying computations of an MLTL formula, many of the functions in the WEST program require the regular expressions of the satisfying computations of the subformulas as inputs. We denote these regexes by $R$ and $T$, where $R = R_1 | R_2 | ... | R_k$ is represented as a set of strings $R = {R_1, R_2, ..., R_k}$, and $T = T_1 | T_2 | ... | T_\ell$ is represented as a set of strings $T = {T_1, T_2, ..., T_\ell}$. Additionally, $n$ will always refer to the number of propositional variables, and nnf refers to an input formula in negation normal form.
\begin{algorithm}[H]
\caption{Pad a set to elements of all equal length}
%Input: set of strings that represents a regex, number of propositional variables \\
%Output: set of strings padded to equal length
\begin{algorithmic}
\Procedure{Pad}{set $R$, int $n$} 
    \State max\_Length $\leftarrow$ $\max_{\{r \in R\}}$($r$.length())
    \For{($r \in R$)}
        \State diff $\leftarrow$ (max\_length$- r$.length()) / $(n+1)$ 
        \State  $r \leftarrow r + (,S^n)^{\text{diff}}$
    \EndFor
    \State return $R$

\EndProcedure
\end{algorithmic}
\end{algorithm}
\begin{algorithm}[H]
\caption{Computes regex for propositional constant}
%Input: String that is either ``T" or ``!", number of propositional variables\\
%Output: set of strings that represents the appropriate satisfying computations
\begin{algorithmic}
\Procedure{reg\_prop\_cons}{string nnf, int $n$}
    \If {(nnf $=$ ``T" and $n \neq 0$)} return $\{S^n\}$
    \Else \hspace{1mm} return $\{\}$ 
    \EndIf
\EndProcedure
\end{algorithmic}
\end{algorithm}
\begin{algorithm}[H]
\caption{Output the set of computation satisfying a propositional variable.}
%Input: String that represents a propositional variable or the negation of one, number of propositional variables \\
%Output: set of strings that represents the appropriate satisfying computations
\begin{algorithmic}
\Procedure{reg\_prop\_var}{string nnf, int $n$}
    \If {(nnf $=$ ``p$k$", where $k$ is a nonnegative integer)} return $\{S^k1S^{n-k-1}\}$
    \EndIf 
    \If {(nnf $=$ ``$\scriptstyle{\sim}$p$k$", where $k$ is a nonnegative integer)} return $\{S^k0S^{n-k-1}\}$
    \EndIf
\EndProcedure
\end{algorithmic}
\end{algorithm}
\begin{algorithm}[H]
\caption{Combines two strings that differ only by one character into one}
%Input: Two strings that represent regexes\\
%Output: Single string that represents computations represented by both input strings or FAIL
\begin{algorithmic}
\Procedure{simplify\_string}{string $r$, string $t$}   
    \If{($r$.length() $\neq$ $t$.length())} exit
        \EndIf
    \For{($0 \leq i < r$.length())}
        \State pre\_r $\leftarrow r[0, i-1]$ 
        \State char\_r $\leftarrow r[i]$ 
        \State post\_r $\leftarrow r[i+1, r\text{.length()}-1]$ 
        \State pre\_t $\leftarrow t[0, i-1]$ 
        \State char\_t $\leftarrow t[i]$ 
        \State post\_t $\leftarrow t[i+1, t\text{.length()}-1]$
        \If{(pre\_r $=$ pre\_t and post\_r $=$ post\_t)}
            \If{(char\_r $\neq$ char\_t)}
                \State return pre\_r +``$S$" + post\_r 
            \Else
                \State return $r$
            \EndIf
        \EndIf
    \EndFor
    \State return FAIL 
\EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
\caption{Simplifies a set of strings using simplify\_string}
%Input: set of strings representing a regex\\
%Output: set of strings representing a regex simplified using simplify\_string
\begin{algorithmic}
\Procedure{simplify}{set $R$}    
    \State Pad all strings in $R$ to the same length
    \If{($R$.length() $\leq 1$)} return $R$
        \EndIf
    \State $i = R\text{.length} - 1$
    \State $j = i - 1$

@@

\caption{Takes union of two regexes (combines two sets into one)}
%Input: sets of strings that represent regexes, simplify boolean\\
%Output: set of strings that represents union of inputted regexes

@@

\Procedure{join}{set $R$, set $T$, bool simp}

@@

% Inputs: set of strings representing the regex for $\phi$, interval bounds, number of propositional variables, simplify boolean\\
% Output: set of strings that represents the appropriate satisfying computations 

@@

\Procedure{reg\_F}{set reg\_phi, int $a$, int $b$, int $n$, bool simp} 

@@

%Inputs: set of strings representing the regex for $\phi$, interval bounds, number of propositional variables, simplify boolean\\
%Output: set of strings that represents the appropriate satisfying computations

@@

\Procedure{reg\_G}{set reg\_phi, int $a$, int $b$, int $n$, bool simp} 

@@

%Inputs: sets of strings representing the regexes for $\phi$ and $\psi$, interval bounds, number of propositional variables, simplify boolean\\
%Output: set of strings that represents the appropriate satisfying computations 
